---
title: "Send R code/R scripts/shell commands to LSF cluster"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Send R code/R scripts/shell commands to LSF cluster}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**Author**: Zuguang Gu ( z.gu@dkfz.de )

**Date**: `r Sys.Date()`

-------------------------------------------------------------


Load the library:

```{r, message = FALSE}
library(bsub)
```

**The _bsub_ package works on the submission nodes, the computing nodes and the Rstudio node.**

**If the current node is not the submission node, a connection via ssh is automatically established.
You might be asked to enter password for the connection.**

## Send R code

The code chunk should be embraced by `{...}`.

```{r, eval = FALSE}
bsub_chunk(name = "example", memory = 10, hour = 10, core = 4, 
{
    fit = NMF::nmf(...)
    # you better save `fit` into a permanent file
    saveRDS(fit, file = "fit.rds")
})
```

In following examples, we use `Sys.sleep(5)` to simulate a chunk of code which runs for a short time.

```{r}
bsub_chunk(
{
	Sys.sleep(5)
})
```

The value returned by `bsub_chunk()` is the job ID from LSF cluster.

### Job settings

Set job name, memory, running time and number of cores:

```{r}
bsub_chunk(name = "example", memory = 10, hour = 10, core = 4, 
{
	Sys.sleep(5)
})
```

If `name` is not specified, an internal name calculated by `digest::digest()` on the chunk
is automatically assigned. The unit of `memory` is GB. 

### R version

The R version can be specified by `R_version` argument.

```{r, eval = FALSE}
bsub_chunk(name = "example", R_version = "3.6.0",
{
	Sys.sleep(5)
})
```

`R_version` can be set as a global parameter:

```{r, eval = FALSE}
bsub_opt$R_version = "3.6.0"
bsub_chunk(name = "example", 
{
	Sys.sleep(5)
})
```

The R version is used for finding proper `Rscript` binary. More generally, `bsub_opt$call_Rscript` is
set as a function which returns the calling of `Rscript`. Also you can initialize other bash environment
variables there. The default is:

```{r, eval = FALSE}
function(version) {
	qq("module load gcc/7.2.0; module load java/1.8.0_131; module load R/@{version}; Rscript")
}
```

where `version` is the value set in `bsub_opt$R_version`. `qq()` is from _GetoptLong_ package which does
[variable interpolation](https://en.wikipedia.org/wiki/String_interpolation).

The module loading for `gcc/7.2.0` and `java/1.8.0_131` ensures that R packages depending on specific C/Java 
libraries can be successfully loaded.

### Bash environment

In previous example, we load the `gcc/7.2.0` and `java/1.8.0_131` modules when loading R. These
bash-level module initialization can also be set by `sh_head` which adds shell commands as header
in the bash script that is used for job submission. E.g. to load R 3.6.0, we can do the other way:

```{r, eval = FALSE}
bsub_opt$call_Rscript = function(version) qq("module load R/@{version}; Rscript")
bsub_chunk(name = "example", sh_head = c("module load gcc/7.2.0", "module load java/1.8.0_131"),
{
	Sys.sleep(5)
})
```

Or set `sh_head` as a global option:

```{r, eval = FALSE}
bsub_opt$call_Rscript = function(version) qq("module load R/@{version}; Rscript")
bsub_opt$sh_head = c("module load gcc/7.2.0", "module load java/1.8.0_131")
bsub_chunk(name = "example",
{
	Sys.sleep(5)
})
```

One usage of this functionality is to load `pandoc` module if the `rmarkdown` is used in the code chunk:

```{r, eval = FALSE}
bsub_chunk(name = "example", sh_head = "module load pandoc/2.2.1",
{
	library(rmarkdown)
	render(...)
})
```

### Load other packages

The packages that are needed can be directly added in the code chunk:

```{r, eval = FALSE}
bsub_chunk(name = "example",
{
	library(package1)
	library(package2)
	Sys.sleep(5)
})
```

Or assign by `packages` argument:

```{r, eval = FALSE}
bsub_chunk(name = "example", packages = c("package1", "package2"),
{
	Sys.sleep(5)
})
```

Or set it as a global parameter:

```{r, eval = FALSE}
bsub_opt$packages = c("package1", "package2")
bsub_chunk(name = "example",
{
	Sys.sleep(5)
})
```

### Other R variables

The R variables that are defined outside the code chunk and need to be used
inside the code chunk can by specified by `variables` argument:

```{r, eval = FALSE}
foo = 1
bsub_chunk(name = "example", variables = "foo",
{ 
	bar = foo
	Sys.sleep(5)
})
```

### The workspace image

If you have too many external variables that are used in the code chunk or
they are used in multiple jobs, you can directly save the workspace or the objects as an
image and specify the `image` argument:

```{r, eval = FALSE}
save.image(file = "/path/foo.RData")
# or 
# save(var1, var2, ..., file = "...")
bsub_chunk(name = "example", image = "/path/foo.RData",
{ 
	...
	Sys.sleep(5)
})
```

Or set the image file as a global parameter:

```{r, eval = FALSE}
save.image(file = "/path/foo.RData")
bsub_opt$image = "/path/foo.RData"
bsub_chunk(name = "example",
{ 
	...
	Sys.sleep(5)
})
```

When `image` is set to `TRUE`, all variables in `.GlobalEnv` will be saved
into a temporary file and all attached packages will be recorded. The temporary files
will be removed after the job is finished.

Absolute paths should be used instead of relative paths.

### The working directory

If the code chunk replies on the working directory, it can be specified by `wd` argument:

```{r, eval = FALSE}
bsub_chunk(name = "example", wd = "/path"
{ 
	Sys.sleep(5)
})
```

Or set it as a global parameter:

```{r, eval = FALSE}
bsub_opt$wd = "/path"
bsub_chunk(name = "example",
{ 
	Sys.sleep(5)
})
```

Absolute path should be used.

### Retrieve the last variable

The last variable in the code chunk can be saved by setting `save_var = TRUE` and retrieved back
by `retrieve_var()` by specifying the job name. Since the variable is looked up by the job name,
there should be no job with the same name submitted before retrieving the variable.

`retrieve_var()` waits until the job is finished.

```{r}
bsub_chunk(name = "example2", save_var = TRUE,
{
	Sys.sleep(10)
	1+1
})
retrieve_var("example2")
```

However, it is not suggested to directly retrieve the returned value from the code chunk. Better choice
is to save the variable into permanent file in the code chunk, E.g.:

```{r, eval = FALSE}
bsub_chunk(name = "example", 
{
	...
	save(...)
	# or
	saveRDS(...)
})
```

### Rerun the job

There is a flag file to mark whether the job was successfully finished or not.
If the job has been successfully done, the job with the same name will be
skipped. `enforce` argument controls how to rerun the jobs with the same
names. If it is set to `TRUE`, jobs will be rerun no matter they are done or not.

```{r, echo = FALSE}
Sys.sleep(10)
```


```{r}
bsub_chunk(name = "example", enforce = FALSE,
{ 
	Sys.sleep(5)
})
```

`enforce` can be set as a global parameter:

```{r, eval = FALSE}
bsub_opt$enforce = FALSE
bsub_chunk(name = "example",
{ 
	Sys.sleep(5)
})
```

### Job dependency

Since `bsub_chunk()` returns the job ID, it is can be used to specify the dependency in other jobs.
The value for `dependency` can be a vector of job IDs.

```{r, eval = FALSE}
job1 = bsub_chunk(name = "example1",
{ 
	Sys.sleep(5)
})
bsub_chunk(name = "example2", dependency = job1,
{ 
	Sys.sleep(5)
})
```

### Temporary and output directory

`bsub_chunk()` has two arguments `temp_dir` and `output_dir`. `temp_dir` is used for the temporary R script
and sh files. `output_dir` is used for the flag files and the output files from LSF cluster.

```{r, eval = FALSE}
bsub_chunk(name = "example", temp_dir = ..., output_dir = ...,
{ 
	Sys.sleep(5)
})
```

They can be set as global parameters. The value of `output_dir` is by default set as the same as `temp_dir`.

```{r, eval = FALSE}
bsub_opt$temp_dir = ...
bsub_opt$output_dir = ...
bsub_chunk(name = "example",
{ 
	Sys.sleep(5)
})
```

To remove temporary files in `temp_dir`, run `clear_temp_dir()` function.

### Run code chunk from a script

You can run code chunk from a script by specifying the starting line number
and the ending line number. The R script is specified by `script` argument,
the starting line number and the ending line number are specified by `start`
and `end` arguments. (Note this functionality has not been tested yet.)

```{r, eval = FALSE}
bsub_chunk(name = "example",
	script = "/path/foo.R",
	start = 10, end = 20, ...)
```

Assume you are editing `foo.R` very offen and the line numbers that you want
to run change from time to time. You can add tags in the R script and
specifying `start` and `end` by those tags. In following example which is the
source code of `foo.R`, we add tags for the code chunk we want to run:

```{r, eval = FALSE}
...
# BSUB_START
you code chunk here
# BSUB_END
...
```

Then you can specify `start` and `end` by regular expressions to match them:

```{r, eval = FALSE}
bsub_chunk(name = "example",
	script = "/path/foo.R",
	start = "^# BSUB_START", 
	end = "^# BSUB_END", ...)
```

### Run jobs locally

Setting `local = TRUE` directly runs the code chunk in the same R session.

```{r}
bsub_chunk(name = "example", local = TRUE,
{ 
	cat("blablabla...\n")
})
```

## Send R script

`bsub_script()` submits the job from R scripts. The major arguments are the same as in `bsub_chunk()`.

```{r, eval = FALSE}
bsub_script("/path/foo.R", name = ..., memory = ..., core = ..., ...)
```

If the R script needs command-line arguments, they can be specified by `argv`.

```{r, eval = FALSE}
bsub_script("/path/foo.R", argv = "--a 1 --b 3", ...)
```

The command-line arguments can also specified as arguments of `bsub_script()`,
but with `.` prefix.

```{r, eval = FALSE}
bsub_script("/path/foo.R", .a = 1, .b = 3, ...)
```

The R scripts should be used with the absolute paths.

Note the bash environment can be initialized by setting a `sh_head`.

## Send other shell commands

`bsub_cmd()`submits shell commands. Basically it is similar as `bsub_script()`:

```{r, eval = FALSE}
bsub_cmd("samtools sort ...", name = ..., memory = ..., core = ..., ...)
bsub_cmd(c("cmd1", "cmd2", ...), name = ..., memory = ..., core = ..., ...)
```

The binary and the arguments should all be set in the first argument of `bsub_cmd()`.

## Job Summary

`bjobs()` or just entering `bjobs` gives a summary of running jobs. Job status (by default is `RUN`
and `PEND`) is controlled by `status` argument. Number of most recent jobs is controlled by `max` argument.
Filtering on the job name is controlled by `filter` argument.

```{r}
library(GetoptLong)
for(i in 1:4) {
	bsub_chunk(name = qq("example_@{i}"),
	{ 
		Sys.sleep(5)
	})
}
bjobs
```

There is one additional column `RECENT` in the summary table which shows the order
of the jobs with the same job name. The most recent job has the value 1.

```{r}
bjobs(status = "all", filter = "example")
```

`brecent()` returns 20 most recent jobs.

```{r}
brecent
```

There are some helper functions which only list running/pending/done/failed jobs:

- `bjobs_running`
- `bjobs_pending`
- `bjobs_done`
- `bjobs_exit`


## Other functions

- `bkill(job_id)` kills a job or a list jobs.
- `job_log(job_id)` prints the log of a specified running/finished/failed job. A vector
  of jobs can also be sent at the same time that last 10 lines of each job are printed.
- `check_dump_files()` searches the dump files (`core.xxx` by LSF cluster or `.RDataTmpxxx` by R).
- `ssh_connect()` establishes the SSH connection to the submission node if it is lost.
- `monitor()` opens a browser-based interactive job monitor.


## Global Parameters

```{r}
bsub_opt
```

## Session Info

```{r}
sessionInfo()
```


