
<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{2. How to use bsub package on different computers}
-->

How to use bsub package on different computers
=====================================

**Author**: Zuguang Gu ( z.gu@dkfz.de )

**Date**: `r Sys.Date()`

-------------------------------------------------------------

## Work with ODCF cluster

### On odcf-worker01

You can submit and monitor jobs by entering commands (under command-line
mode), but you cannot use the interactive job monitor (because web browser is
not available on `odcf-worker01`).

### On B080/W610 pc

You will be asked for the password for connecting to the submission node via
ssh. Don't submit the jobs there because the file system is different from
that on the submission nodes. You can open the interactive job monitor by
`monitor()`. You might have the following problem saying you cannot open the
web browser:

```
monitor()
# Listening on http://127.0.0.1:5298
# Couldn't get a file descriptor referring to the console
```

You can directly copy `http://127.0.0.1:5298` (the port can be different for
you) and paste it to the web browser, or you need to set the global option for
finding the browser:

```{r, eval = FALSE}
options("browser" = Sys.which("firefox"))
monitor()  # now firefox will be automatically opened.
```

### On your own laptop

You need to connect VPN first, and then load the ODCF configuration:

```{r, eval = FALSE}
bsub:::config_odcf(your_user_name)
```

After that you can use the interactive job monitor.

### On Rstudio server

You can submit jobs, and also monitor your jobs with the interactive job monitor.

## Work for other institutes

Please see [the last section in the main vignette](bsub_intro.html).
