---
title: "Configure bsub package for other LSF institutes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3. Configure bsub package for other LSF institutes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**Author**: Zuguang Gu ( z.gu@dkfz.de )

**Date**: `r Sys.Date()`

-------------------------------------------------------------

The **bsub** package is only broadly tested on DKFZ ODCF cluster and the
default configurations work fine there. It is also possible for other clusters
that use LSF system to use **bsub** package. Following global options should
be properly configured.

### How to call Rscript binary

You need to configure how to call the `Rscript` binary, especially when you
have different versions of R installed. Following is how we call `Rscript` on
our cluster (we use `module` to manage the bash environment).

```{r, eval = FALSE}
bsub_opt$call_Rscript = function(version) qq("module load gcc/7.2.0; module load java/1.8.0_131; module load R/@{version}; Rscript")
```

Then the different version of `Rscript` can be switched by setting different
`bsub_opt$R_version` option.


### Submission node

The names of the nodes where the jobs are submitted can be set by
`bsub_opt$submission_node` option. The value can be a vector of node names.
**bsub** will randomly select one to connect.

```{r, eval = FALSE}
bsub_opt$submission_node = ...
```

### Username on the submission node

Your username on the submission node can be set by `bsub_opt$user`. The
default value is `Sys.info()['user']`. If the username on the submission node
is different, you need to explicitly set it.

```{r, eval = FALSE}
bsub_opt$user = ...
```

### Bash environment

`bsub_opt$ssh_envir` should be properly set so that LSF binaries such as
`bsub` or `bjobs` can be properly found. There are some environment variables
initialized when logging in the bash terminal while they are not initialized
with the ssh connection. Thus, some environment variables should be manually
set.

An example for `bsub_opt$ssh_envir` is as follows. The `LSF_ENVDIR` and
`LSF_SERVERDIR` should be defined and exported.
 
```{r, eval = FALSE}
bsub_opt$ssh_envir = c("source /etc/profile",
                       "export LSF_ENVDIR=/opt/lsf/conf",
                       "export LSF_SERVERDIR=/opt/lsf/10.1/linux3.10-glibc2.17-x86_64/etc")
```


The values of these two variables can be obtained by entering following
commands in your bash terminal (on the submission node):

```
echo $LSF_ENVDIR
echo $LSF_SERVERDIR
```

### bsub template

You need to define the template for calling the `bsub` command by
`bsub_opt$bsub_template` option. The self-defined function should accepts
following arguments:

- `name` job name.
- `hour` running time.
- `memory` memory, in GB.
- `core` number of cores to use.
- `output` path of output file.
- `...` should be added as the last argument of the function. Following is a
  simple example for calling `bsub`.

E.g., the template on DKFZ ODCF is defined as:

```{r, eval = FALSE}
bsub_opt$bsub_template = function(name, hour, memory, core, output, ...) {
	glue::glue("bsub -J '{name}' -W '{hour}:00' -n {core} -R 'rusage[mem={memory}GB]' -o '{output}'")
}
```

You can use `glue::glue()` or `GetoptLong::qq()` to construct a complex string
with multiple variables.

### How to parse the time strings

The time strings by LSF `bjobs` command might be different for different
configurations. The **bsub** package needs to convert the time strings to
`POSIXct` objects for calculating the time difference. Thus, if the default
time string parsing fails, users need to provide a user-defined function and
set with `bsub_opt$parse_time` option in `bsub_opt`. The function accepts a
vector of time strings and returns a `POSIXct` object. For example, if the
time string returned from `bjobs` command is in a form of `Dec 1 18:00:00
2019`, the parsing function can be defined as:

```{r, eval = FALSE}
bsub_opt$parse_time = function(x) {
    as.POSIXct(x, format = "%b %d %H:%M:%S %Y")
}
```


